%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{xcolor}
\usepackage{listings}
\usepackage{graphicx}


\usepackage{txfonts}
%\usepackage{newtxtext}       %
%\usepackage{newtxmath}       % selects Times Roman as basic font

%\usepackage{subcaption}
\usepackage{multirow}
\usepackage{wrapfig}

\newenvironment{tightItemize}{
        \begin{itemize}
                \setlength{\itemsep}{1pt}
                \setlength{\parskip}{0pt}
                \setlength{\parsep}{0pt}
        }{\end{itemize}
}
\newcommand{\fix}[1]{\textcolor{red}{#1}} %Put words in Red

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title*{Ascent: A Flyweight In Situ Library for Exascale Simulations}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Matthew Larsen, Eric Brugger, Hank Childs, and Cyrus Harrison}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Matthew Larsen, Eric Brugger, and Cyrus Harrison \at Lawrence Livermore National Laboratory
\and Hank Childs \at University of Oregon}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle




%\chapter{Ascent: A Flyweight In Situ Library for Exascale Simulations}
\label{ascent} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head

\abstract{
This chapter describes Ascent, a production library for in situ visualization and analysis
on exascale architectures.
%
It begins by describing the library's focal points:
minimizing encumbrance on simulation codes and enabling diverse
and powerful capabilities.
%
The chapter then describes Ascent's abstractions, design, and interface.
%
It concludes with success stories that  highlight its
capabilities:
in situ visualization of a 97.8 billion element intertial confinement fusion
simulation using 16,384 GPUs,
delivering radiography capabilities for a Kelvin-Helmholtz simulation,
and native rendering of higher order elements.
}

\section{Introduction}

This chapter describes the Ascent library for in situ visualization
and analysis.
%
Ascent was ``born in situ,'' meaning that it was created for the
express purpose of performing in situ processing, and all implementation
decisions follow this mandate.
%
It was designed to deliver on two goals which are often in tension:
 (1) minimizing encumbrance on simulation codes that incorporate Ascent,
i.e., a ``flyweight'' design (discussed in \S\ref{sec:flyweight}), and
(2) providing diverse and powerful capabilities,
especially for modern supercomputers (discussed in \S\ref{sec:subcapabilities}).
%
In terms of the taxonomy presented in \fix{Intro chapter, section 3}:
\begin{tightItemize}
\item Ascent has an API for visualization and analysis that is directly
used by a computational simulation.  It is primarily controlled by
specifying the desired visualization and analyses a priori, but it
also has an option to perform human-in-the-loop via Jupyter notebooks.
\item Ascent can be run in multiple ways, but the primary usage is for
Ascent to share the same resources as the simulation code (space division),
with direct access to the simulation code's memory.
\end{tightItemize}
Finally,  while not a focus of this chapter, Ascent is built for production use by computational
simulation codes --- it has documentation,
examples, engages in modern software engineering practices, etc.

\subsection{Flyweight Design}
\label{sec:flyweight}

Since Ascent was designed to run on the same resources as a simulation code,
it is critical that it minimizes its usage of those resources.
%
In particular, Ascent aims to minimize
execution time, memory usage, binary size, and integration effort.

The first two aims, minimizing execution time and memory usage, benefit from directly
incorporating VTK-m~\cite{Moreland:CGA2016} into Ascent.
%
For minimizing memory,
VTK-m's data model supports many array layouts (e.g., row major versus
column major, array of structures versus structures of arrays)
in a zero-copy manner.
%
In most cases, the majority of the memory needed by Ascent is only for
storing intermediate results, such as the triangles produced by an isosurface algorithm.
%
To minimize execution time,
VTK-m provides native support for many-core architectures,
including NVIDIA GPUs, various Intel architectures,
and planned support for AMD GPUs.
%
Further, Ascent complements VTK-m's shared-memory parallelism with
its own layer for distributed-memory parallelism based on the
Message Passing Interface (MPI) library.
%
In all, Ascent is able to perform its algorithms very quickly,
since it can make full use of the underlying hardware architecture.
%
%\fix{Hank's version:}
%In terms of binary size, Ascent has a focus on minimizing dependencies on other libraries.
%%
%In addition to a small binary size, this also allows Ascent to be ported to new supercomputers
%more quickly.
%%
%For example, Ascent can make use of OpenGL, but it also has a software-only renderer that allows
%it to proceed without OpenGL.
%%
%In all, dependencies are only brought in as needed.
%%
%Further, to integrate Ascent, simulation codes only need to work with Ascent's API.
%%
%\fix{All other symbols are imported by Ascent during run-time.}
%\fix{Matt's version:}
%Ascent also has far fewer dependencies than more mature community
%visualization tools which leads to smaller binary sizes.

In terms of binary size, Ascent has far
fewer dependencies than more mature community visualization tools.
%
Not only does this reduce build complexity, but it results in
smaller binary sizes, lessening the linking burden on the simulation.
%
The only required libraries for Ascent are VTK-m
and Conduit (described in \S\ref{Conduit}).
%
Ascent also adds its own libraries, specifically VTK-h (an MPI wrapping around
VTK-m) and Flow (described in \S\ref{sec:flow}).
%
All other libraries are optional.

Finally, Ascent's API was designed to be simple to learn,
easy to integrate for novice users, and small in overall code size.
%
Integrations can be as short as ten lines of code.
%
This topic is explored more in \S\ref{sec:API}.

\subsection{Ascent Capabilities}
\label{subsec:capabilities}

Ascent takes multiple approaches to deliver diverse and powerful capabilities:
integrating with many technologies,
supporting both visualization and analysis routines,
developing algorithms specifically designed for in situ processing,
and providing support for modern supercomputers.

Ascent's strongest component for delivering capabilities is likely
through integrating many additional technologies.
%
In terms of integrations with other visualization and I/O products,
Ascent is able to produce Cinema and HDF5 files, interact with
ADIOS and Catalyst, and, as previously mentioned, utilizes VTK-m
directly for much of its data model and visualization algorithms.
%
It also provides many options for integration, including
bindings
for controlling Ascent with  C, C++, Python, and Fortan.
%
Further, it provides options for controlling Ascent via its code API,
YAML files (which can be changed dynamically at runtime), or
a Jupyter notebook integration \cite{CyrusISAV}.
%
Finally, Ascent is able to integrate arbitrary code, including
Python scripts and C/C++ code.
%
This pathway also enables access to
machine learning infrastructures, such as TensorFlow and PyTorch.
%
This is more than proof of concept,
as Ascent recently utilized a distributed-memory Random Forest.
%
A successful integration of code is described in \S\ref{sec:DevilRay}.
%
Specifically, this integration incorporated the Devil Ray code for ray-tracing
higher order elements.
%\S\ref{sec:DevilRay} describes a success story regarding integrating\S\ref{sec:DevilRay}
%code,
%namely incorporating the Devil Ray code for ray-tracing higher order
%elements.

Ascent has embraced analysis alongside visualization.
%
Noteworthy analysis examples include simulated radiographs and its
query infrastructure to get quantitative results such as volumes,
surfaces areas, and integrated quantities (e.g., integrate the density field to get the mass).
%
Of course, Ascent supports typical visualization algorithms, such as
slicing, isosurfacing, volume rendering, and multiple plot types.
%
With respect to algorithms specifically for in situ processing, Ascent
has a significant investment in its ``trigger'' system~\cite{Larsen:ISAV18},
which adapts when visualization and analysis routines are executed,
based on the conditions in the system.
%
Ascent also has algorithms for extracting reduced forms of the data,
such as a module for extracting Lagrangian flow~\cite{Agranovsky:LDAV2014,Sane:EGPGV19} and the
aforementioned Cinema output.

Finally, Ascent has been designed from the beginning with modern
supercomputers in mind.
%
Especially because of the integration of VTK-m, Ascent is able to
run efficiently on these architectures.

\subsection{Organization of This Chapter}

The remainder of this chapter focuses on key elements of Ascent:
\begin{itemize}
\item \S\ref{sec:capabilities} discusses the five abstractions for using Ascent.
\item \S\ref{sec:design} discusses Ascent's design.
\item \S\ref{sec:API} discusses Ascent's APIs for inputting
data and for controlling execution.
\item \S\ref{sec:success} discusses success stories using Ascent.
\item \S\ref{sec:information} discusses where to find more infomation about Ascent.
\end{itemize}

\section{Key Abstractions for Ascent}
\label{sec:capabilities}
\input{ascent_does_stuff}

\section{System Architecture}
\label{sec:design}
\input{system_architecture}

\section{Ascent APIs}
\label{sec:API}
\input{api}

\section{Success Stories}
\label{sec:success}
\input{successes}

\subsection{Devil Ray Rendering}
\label{sec:DevilRay}
\input{devil_ray}

\section{Additional Resources}
\label{sec:information}
\input{information}
%\fix{EVERYTHING BELOW THIS POINT IS CURRENTLY ORPHANED.  AUTHORS
%SHOULD PULL INFORMATION FROM THE BELOW AS THEY POPULATE THE SECTIONS
%ABOVE.}


\if 0
Statement about capabilities.  Emphasis on being an integrator.
This is where Jupyter, triggers, etc. get mentioned.  Also emphasize
exascale.
Ascent has a rich feature set and integrates with other ecosystems.

\begin{itemize}
  \item Supports common visualization operations such as slicing and dicing your data.
  \item Supports ray traced surface and volume rendering as well as simulated radiography.
  \item Supports Queries for getting quantitative answers to questions
  \item Supports Triggers for adaptive visualization and analysis
  \item Outputs images sequences and Cinema image databases
  \item Outputs extracts for exporting data to HDF5, ADIOS, Python and Jupyter
  \item Support interactive visualization and analysis with Jupyter and Catalyst
  \item Supports C/C++, Python and Fortran language bindings
\end{itemize}

Statement about Devil Ray, but a forward reference a section in success stories --- success as an integrator.
\fi


%\section{The Ascent In Situ Infrastructure}
%\label{ascent_overview}
%Use the template \emph{chapter.tex} together with the document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content conformable to the Springer Nature layout.

%List how ascent makes things easy.

%\section{Design Considerations}
%\label{ascent_design_considerations}
%\input{design_considerations}


%\section{Data Interface}
%\label{ascent_data}
%\input{data_interface}

%\section{Control Interface}
%\label{ascent_control}
%\input{control_interface}

%\section{Success Stories}
%\label{ascent_success}

\bibliographystyle{splncs04}
\bibliography{book}
\end{document}
