This subsection describes the basics of Conduit, a library for describing 
hierarchical scientific data in C++, C, Fortran, and Python.
%
%that provides an intuitive model 
The API was designed to be intuitive, and is inspired by JSON.
%
While primarily used for data coupling between packages in-core,
Conduit also provides easy access to serialization and I/O functions.
%
%
%In this section, we will discuss some of the basics of Conduit
%and describe the concepts needed to couple simulation mesh
%data structures using Conduit's Mesh Blueprint.

The primary data structure in Conduit is a Node.
%
Nodes stores data through a key-value interface.
%
Fig.~\ref{ex:1} shows how to store the string ``value'' with the key ``K'' into a Node.

\begin{figure}
\begin{tabular}{cc}
  \begin{minipage}{.5\textwidth}
  \centering
    \begin{lstlisting}[language=C++]
Node n;
n["K"] = "value";
n.print();
    \end{lstlisting}
  \end{minipage}
  &
  \begin{minipage}{.5\textwidth}
  \centering
  \begin{lstlisting}[language=C++]
K: "value"
  \end{lstlisting}
  \end{minipage}
\end{tabular}
\caption{\label{ex:1}On the left, an example of storing a string inside a Conduit Node. On the right, the YAML equivalent.  The YAML has increasing indentation to indicate increasing depth in a hierarchy.}
\end{figure}

Data in Nodes can be created and accessed through hierarchical keys,
and the key string looks much like a UNIX directory structure.
%
Key paths are completely up to the user.
%
Fig.~\ref{ex:2} illustrates creating a Node hierarchy
and storing a floating point value in a leaf Node.
%
In this example, several Nodes are actually created with parent child
relationships defined by the key.
%
Node $n$ has a child Node $dir1$, which in turn has a child $dir2$,
and the tree ends at the leaf Node $val1$, which stores the data.
%
Nodes can contain many basic data types such as strings,
floating point values, and integers.

\begin{figure}
\begin{tabular}{cc}
  \begin{minipage}{.5\textwidth}
  \centering
    \begin{lstlisting}[language=C++]
Node n;
n["dir1/dir2/val1"] = 100.5;
n.print();
    \end{lstlisting}
  \end{minipage}
  &
  \begin{minipage}{.5\textwidth}
  \centering
  \begin{lstlisting}[language=C++]
dir1:
  dir2:
    val1: 100.5
  \end{lstlisting}
  \end{minipage}
\end{tabular}
\caption{\label{ex:2}On the left, an example of using a hierarchical key to store a number. On the right, the YAML equivalent.}
\end{figure}

Nodes can also contain arrays, and the $set$ method
copies the values from an array into a Node.
%
Fig.~\ref{ex:3} shows how to set a Node to the contents of an array.
%
Alternatively, the $set\_external$ method copies only the pointer (i.e., a shallow copy),
and any change to the underlying array would be reflected in both the original
array and the data contained inside the Node.
%
Using $set\_external$ is desirable for large data, such as simulation mesh data,
and in environments where memory is constrained.

\begin{figure}
\begin{tabular}{cc}
  \begin{minipage}{.5\textwidth}
  \centering
    \begin{lstlisting}[language=C++]
const int size = 4;
int A[size] = {0, 1, 2, 3};
Node n;
n["my_array"].set(A, size);
n.print();
    \end{lstlisting}
  \end{minipage}
  &
  \begin{minipage}{.5\textwidth}
  \centering
  \begin{lstlisting}[language=C++]
my_array: [0, 1, 2, 3]
  \end{lstlisting}
  \end{minipage}
\end{tabular}

\caption{\label{ex:3}On the left, an example of setting the value of a Node to an Array. On the right, the YAML equivalent.}
\end{figure}

As already mentioned, Conduit lays the foundation for Ascent's API.
%
That said, Conduit enjoys widespread use outside Ascent.
%
It is being used for simulation code coupling, effectively as a data store, among
other use cases.

%%\fix{At the end, we should mention how simulations codes are using conduit as
%a central data store, and for checkpoints. It also might be nice to
%illustrate some other interesting use cases like neurons, or diffing nodes
%to update rendering states in wolf vision. The limits of using using Conduit
%are only bounded by the imagination of the person who wields Conduit.}

